---
title: "Apertus validity — primary analysis"
subtitle: "Eight TOSTs, conjunction, Bland–Altman, and the sampling-frame ledger"
format:
  html:
    toc: true
    number-sections: true
    embed-resources: true
execute:
  echo: true
  warning: false
params:
  evaluation: "../fixtures/evaluation.csv"
  ledger: "../fixtures/ledger.csv"
---

# What this document answers

The pre-registered design is
[`docs/framework-empirical-evaluation.md`](../../docs/framework-empirical-evaluation.md).
This Quarto document is the primary analysis, not an exploratory notebook.

The estimand is the mean relative error on the executable Apertus-like grid.
A **subject** is one architecture together with one \((B, S)\). Each subject
contributes one relative error to each of eight cells (phase × precision ×
hypothesis). Extra training steps are not extra \(n\).

| Hypothesis | Zepto column \(y\) | Twin column \(t\) |
|------------|--------------------|-------------------|
| H1 VRAM | `y_vram` | `target_vram` |
| H2 FLOPs | `y_flop_fcm` | `target_flop` |

Relative error:

\[
\mathrm{RE} = \frac{y - t}{t}.
\]

Equivalence margin \(\Delta = 0.10\). \(\alpha = 0.05\) for each pair of
one-sided tests. No Bonferroni. Family validity is the **conjunction** of
all eight TOSTs. An inconclusive TOST is not evidence that Zepto is
inaccurate to ten percent.

`y_flop` is stored on the CSV as a diagnostic. It is not the H2 column.

```{r}
#| label: setup
required <- c("TOSTER")
missing <- required[!vapply(required, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing) > 0) {
  stop("Install missing R packages: ", paste(missing, collapse = ", "),
       "\n  install.packages(c(\"TOSTER\"))", call. = FALSE)
}

relative_error <- function(y, t) {
  if (any(t == 0, na.rm = TRUE)) {
    stop("twin probe t must be non-zero", call. = FALSE)
  }
  (y - t) / t
}

MARGIN <- 0.10
ALPHA <- 0.05

evaluation <- read.csv(params$evaluation, stringsAsFactors = FALSE)
ledger_path <- params$ledger
ledger <- if (file.exists(ledger_path)) {
  read.csv(ledger_path, stringsAsFactors = FALSE)
} else {
  data.frame()
}

needed <- c(
  "subject_id", "phase", "precision",
  "y_vram", "target_vram", "y_flop_fcm", "target_flop"
)
missing_cols <- setdiff(needed, names(evaluation))
if (length(missing_cols) > 0) {
  stop("evaluation.csv missing: ", paste(missing_cols, collapse = ", "),
       call. = FALSE)
}

evaluation$re_vram <- relative_error(evaluation$y_vram, evaluation$target_vram)
evaluation$re_flop <- relative_error(evaluation$y_flop_fcm, evaluation$target_flop)
```

# Data

Primary analysis uses **completed subjects only**. The ledger is the
sampling-frame appendix (framework §5.3).

```{r}
#| label: data-counts
n_rows <- nrow(evaluation)
n_subjects <- length(unique(evaluation$subject_id))
n_ledger <- if (nrow(ledger) == 0) 0 else nrow(ledger)

knitr::kable(
  data.frame(
    evaluation_rows = n_rows,
    completed_subject_ids = n_subjects,
    ledger_rows = n_ledger
  ),
  caption = "Completed measurements versus ledger rows."
)
```

# Eight one-sample TOSTs

For each cell the tests are (framework §5.1):

- \(H_{01}\colon \mu \le -0.10\) versus \(H_{11}\colon \mu > -0.10\)
- \(H_{02}\colon \mu \ge +0.10\) versus \(H_{12}\colon \mu < +0.10\)

Equivalence is declared only if both one-sided tests reject. The two
`t.test` calls below are that definition.

`TOSTER::t_TOST` is the same two one-sided tests packaged by Lakens and
Caldwell. It is confirmation only; it does not decide the cell. H2
inference can have **zero** relative-error SD: every subject’s
`y_flop_fcm` equals `target_flop`. That is a valid TOST input
(\(t = \pm\infty\)). TOSTER 0.8.6 still tries to invert a non-central
\(t\) for a Cohen’s \(d\) interval (`get_ncp_t2`) and errors with
`object 'Best.Low' not found`. Confirmation is skipped when the SD is
zero.

```{r}
#| label: tost-cells
cells <- expand.grid(
  hypothesis = c("vram", "flop"),
  phase = c("inference", "training"),
  precision = c("fp16", "fp32"),
  stringsAsFactors = FALSE
)

tost_one_cell <- function(hypothesis, phase, precision) {
  part <- evaluation[
    evaluation$phase == phase & evaluation$precision == precision,
  ]
  part <- part[!duplicated(part$subject_id), ]
  re <- if (hypothesis == "vram") part$re_vram else part$re_flop
  re <- re[is.finite(re)]
  n <- length(re)
  if (n < 2) {
    return(data.frame(
      hypothesis = hypothesis,
      phase = phase,
      precision = precision,
      n = n,
      mean_re = NA_real_,
      sd_re = NA_real_,
      p_lower = NA_real_,
      p_upper = NA_real_,
      equivalent = FALSE,
      decision = "not_enough_n",
      stringsAsFactors = FALSE
    ))
  }

  # Written-out two one-sided tests (Schuirmann 1987).
  # Zero SD is allowed: all observations equal, t is ±Inf.
  lower <- t.test(re, alternative = "greater", mu = -MARGIN)
  upper <- t.test(re, alternative = "less", mu = MARGIN)
  both_reject <- (lower$p.value < ALPHA) && (upper$p.value < ALPHA)

  # Packaged confirmation (Lakens 2017). Skip when SD is 0: TOSTER's
  # Cohen's d CI helper does not accept an infinite t-statistic.
  sd_re <- stats::sd(re)
  if (is.finite(sd_re) && sd_re > 0) {
    invisible(TOSTER::t_TOST(
      x = re,
      hypothesis = "EQU",
      eqb = MARGIN,
      eqbound_type = "raw",
      alpha = ALPHA
    ))
  }

  decision <- if (both_reject) {
    "equivalent"
  } else {
    "inconclusive_or_not_equivalent"
  }

  data.frame(
    hypothesis = hypothesis,
    phase = phase,
    precision = precision,
    n = n,
    mean_re = mean(re),
    sd_re = sd_re,
    p_lower = unname(lower$p.value),
    p_upper = unname(upper$p.value),
    equivalent = both_reject,
    decision = decision,
    stringsAsFactors = FALSE
  )
}

tost_table <- do.call(
  rbind,
  Map(tost_one_cell, cells$hypothesis, cells$phase, cells$precision)
)
rownames(tost_table) <- NULL

knitr::kable(
  tost_table[, c(
    "hypothesis", "phase", "precision", "n",
    "mean_re", "sd_re", "p_lower", "p_upper", "equivalent", "decision"
  )],
  digits = 4,
  caption = "One-sample TOST per phase × precision × hypothesis. p_lower is H01; p_upper is H02."
)
```

# Conjunction

The family claim holds only if every cell above is `equivalent`.

```{r}
#| label: conjunction
family_pass <- all(tost_table$equivalent)
cat(
  if (family_pass) {
    "Conjunction: PASS. All eight TOSTs declared equivalence."
  } else {
    paste(
      "Conjunction: FAIL.",
      "At least one cell is not equivalent or is inconclusive.",
      "An inconclusive cell is not evidence that Zepto is inaccurate to 10%."
    )
  },
  "\n"
)
knitr::kable(
  tost_table[!tost_table$equivalent, c("hypothesis", "phase", "precision", "decision")],
  caption = "Cells that did not declare equivalence (empty if the conjunction passed)."
)
```

# Bland–Altman

Framework §5.2. These figures do not replace TOST. Horizontal axis
\((y + t) / 2\); vertical axis signed difference \(y - t\). Each panel
shows the mean difference and the limits of agreement (mean ± 1.96 SD).

```{r}
#| label: bland-altman
#| fig-width: 9
#| fig-height: 8
old_par <- par(mfrow = c(4, 2), mar = c(4, 4, 3, 1))
on.exit(par(old_par), add = TRUE)

for (i in seq_len(nrow(cells))) {
  hyp <- cells$hypothesis[[i]]
  phase <- cells$phase[[i]]
  precision <- cells$precision[[i]]
  part <- evaluation[
    evaluation$phase == phase & evaluation$precision == precision,
  ]
  part <- part[!duplicated(part$subject_id), ]
  y <- if (hyp == "vram") part$y_vram else part$y_flop_fcm
  t <- if (hyp == "vram") part$target_vram else part$target_flop
  average <- (y + t) / 2
  difference <- y - t
  mean_diff <- mean(difference)
  loa <- 1.96 * sd(difference)

  plot(
    average,
    difference,
    xlab = "(y + t) / 2",
    ylab = "y - t",
    main = paste(hyp, phase, precision),
    pch = 16
  )
  abline(h = 0, lty = 3)
  abline(h = mean_diff, lty = 1)
  abline(h = mean_diff + loa, lty = 2)
  abline(h = mean_diff - loa, lty = 2)
}
```

# Ledger (sampling-frame appendix)

Builds that fail, including out-of-memory failures, are not dropped
silently. They are not in the eight TOSTs.

```{r}
#| label: ledger
if (nrow(ledger) == 0) {
  cat("No ledger rows.\n")
} else {
  knitr::kable(ledger, caption = "Failed attempts. Primary analysis uses completed subjects only.")
}
```
