The estimand is the mean relative error on the executable family-like grid.
A **subject** is one architecture together with one $(B, S)$. Each
subject contributes one relative error to each of eight cells formed by the
crossing of phase, precision, and hypothesis. Additional training steps do not
increase the sample size $n$.

| Hypothesis | Zepto column $y$ | Twin column $t$ |
|------------|------------------|-----------------|
| H1 VRAM | `y_vram` | `target_vram` |
| H2 FLOPs | `y_flop_fcm` | `target_flop` |

: Columns that enter each hypothesis as the Zepto prediction $y$ and the twin probe $t$. {#tbl-hypotheses}

The relative error is defined in @eq-relative-error.

$$
\mathrm{RE} = \frac{y - t}{t}
$$ {#eq-relative-error}

The equivalence margin is $\Delta = 0.10$ and the significance level is
$\alpha = 0.05$ for each pair of one sided tests. Multiplicity is not adjusted
with a Bonferroni correction. Family validity is the **conjunction** of all
eight TOSTs, so the family claim is supported only when every cell declares
equivalence. An inconclusive TOST does not constitute evidence that Zepto is
inaccurate to ten percent.

The column `y_flop` is stored on the CSV as a diagnostic and is not the column
used for H2.

The following chunk loads TOSTER, defines relative error, and reads the
evaluation table together with the ledger.

```{r}
#| label: setup
required <- c("TOSTER")
missing <- required[!vapply(required, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing) > 0) {
  stop("Install missing R packages: ", paste(missing, collapse = ", "),
       "\n  install.packages(c(\"TOSTER\"))", call. = FALSE)
}

relative_error <- function(y, t) {
  if (any(t == 0, na.rm = TRUE)) {
    stop("twin probe t must be non-zero", call. = FALSE)
  }
  (y - t) / t
}

MARGIN <- 0.10
ALPHA <- 0.05

evaluation <- read.csv(params$evaluation, stringsAsFactors = FALSE)
ledger_path <- params$ledger
ledger <- if (file.exists(ledger_path)) {
  read.csv(ledger_path, stringsAsFactors = FALSE)
} else {
  data.frame()
}

needed <- c(
  "subject_id", "phase", "precision",
  "y_vram", "target_vram", "y_flop_fcm", "target_flop"
)
missing_cols <- setdiff(needed, names(evaluation))
if (length(missing_cols) > 0) {
  stop("evaluation.csv missing: ", paste(missing_cols, collapse = ", "),
       call. = FALSE)
}

evaluation$re_vram <- relative_error(evaluation$y_vram, evaluation$target_vram)
evaluation$re_flop <- relative_error(evaluation$y_flop_fcm, evaluation$target_flop)
```

# Data {#sec-data}

The primary analysis uses **completed subjects only**. The ledger is retained as
the sampling frame appendix described in framework section 5.3.
@tbl-data-counts reports how many evaluation rows and completed subject
identifiers are present, together with the number of ledger rows.

```{r}
#| label: tbl-data-counts
#| tbl-cap: "Completed measurements versus ledger rows."
n_rows <- nrow(evaluation)
n_subjects <- length(unique(evaluation$subject_id))
n_ledger <- if (nrow(ledger) == 0) 0 else nrow(ledger)

knitr::kable(
  data.frame(
    evaluation_rows = n_rows,
    completed_subject_ids = n_subjects,
    ledger_rows = n_ledger
  )
)
```

# Eight one sample TOSTs {#sec-tost}

For each cell the tests follow framework section 5.1. The first one sided test
is $H_{01}\colon \mu \le -0.10$ versus $H_{11}\colon \mu > -0.10$, and the
second is $H_{02}\colon \mu \ge +0.10$ versus $H_{12}\colon \mu < +0.10$.
Equivalence is declared only if both one sided tests reject, and the two
`t.test` calls below implement that definition.

`TOSTER::t_TOST` packages the same two one sided tests following Lakens and
Caldwell. It is used only as confirmation and does not decide the cell. H2
inference can have a relative error standard deviation of zero when every
subject's `y_flop_fcm` equals `target_flop`. That configuration remains a valid
TOST input, with $t = \pm\infty$. TOSTER 0.8.6 still tries to invert a
noncentral $t$ for a Cohen's $d$ interval through `get_ncp_t2` and then errors
with `object 'Best.Low' not found`. Confirmation is therefore skipped when the
standard deviation is zero. @tbl-tost-cells reports the eight cells.

```{r}
#| label: tbl-tost-cells
#| tbl-cap: "One sample TOST per phase, precision, and hypothesis. p_lower is H01 and p_upper is H02."
cells <- expand.grid(
  hypothesis = c("vram", "flop"),
  phase = c("inference", "training"),
  precision = c("fp16", "fp32"),
  stringsAsFactors = FALSE
)

tost_one_cell <- function(hypothesis, phase, precision) {
  part <- evaluation[
    evaluation$phase == phase & evaluation$precision == precision,
  ]
  part <- part[!duplicated(part$subject_id), ]
  re <- if (hypothesis == "vram") part$re_vram else part$re_flop
  re <- re[is.finite(re)]
  n <- length(re)
  if (n < 2) {
    return(data.frame(
      hypothesis = hypothesis,
      phase = phase,
      precision = precision,
      n = n,
      mean_re = NA_real_,
      sd_re = NA_real_,
      p_lower = NA_real_,
      p_upper = NA_real_,
      equivalent = FALSE,
      decision = "not_enough_n",
      stringsAsFactors = FALSE
    ))
  }

  # Written-out two one-sided tests (Schuirmann 1987).
  # Zero SD is allowed: all observations equal, t is ±Inf.
  lower <- t.test(re, alternative = "greater", mu = -MARGIN)
  upper <- t.test(re, alternative = "less", mu = MARGIN)
  both_reject <- (lower$p.value < ALPHA) && (upper$p.value < ALPHA)

  # Packaged confirmation (Lakens 2017). Skip when SD is 0: TOSTER's
  # Cohen's d CI helper does not accept an infinite t-statistic.
  sd_re <- stats::sd(re)
  if (is.finite(sd_re) && sd_re > 0) {
    invisible(TOSTER::t_TOST(
      x = re,
      hypothesis = "EQU",
      eqb = MARGIN,
      eqbound_type = "raw",
      alpha = ALPHA
    ))
  }

  decision <- if (both_reject) {
    "equivalent"
  } else {
    "inconclusive_or_not_equivalent"
  }

  data.frame(
    hypothesis = hypothesis,
    phase = phase,
    precision = precision,
    n = n,
    mean_re = mean(re),
    sd_re = sd_re,
    p_lower = unname(lower$p.value),
    p_upper = unname(upper$p.value),
    equivalent = both_reject,
    decision = decision,
    stringsAsFactors = FALSE
  )
}

tost_table <- do.call(
  rbind,
  Map(tost_one_cell, cells$hypothesis, cells$phase, cells$precision)
)
rownames(tost_table) <- NULL

knitr::kable(
  tost_table[, c(
    "hypothesis", "phase", "precision", "n",
    "mean_re", "sd_re", "p_lower", "p_upper", "equivalent", "decision"
  )],
  digits = 4
)
```

# Conjunction {#sec-conjunction}

The family claim holds only if every cell in @tbl-tost-cells is marked
`equivalent`. The following chunk states whether that conjunction holds and
lists any cell that did not declare equivalence.

```{r}
#| label: conjunction
family_pass <- all(tost_table$equivalent)
cat(
  if (family_pass) {
    "Conjunction: PASS. All eight TOSTs declared equivalence."
  } else {
    paste(
      "Conjunction: FAIL.",
      "At least one cell is not equivalent or is inconclusive.",
      "An inconclusive cell is not evidence that Zepto is inaccurate to 10%."
    )
  },
  "\n"
)
knitr::kable(
  tost_table[!tost_table$equivalent, c("hypothesis", "phase", "precision", "decision")],
  caption = "Cells that did not declare equivalence (empty if the conjunction passed)."
)
```

# Bland Altman {#sec-bland-altman}

Framework section 5.2 asks for Bland Altman figures as a descriptive companion,
and these figures do not replace TOST. @fig-bland-altman plots the pairwise
average $(y + t) / 2$ on the horizontal axis and the signed difference $y - t$
on the vertical axis. Each panel shows the mean difference and the limits of
agreement, computed as the mean plus or minus 1.96 standard deviations.

```{r}
#| label: fig-bland-altman
#| fig-cap: "Bland Altman panels for each phase, precision, and hypothesis. The horizontal axis is the pairwise average and the vertical axis is the signed difference."
#| fig-width: 9
#| fig-height: 8
old_par <- par(mfrow = c(4, 2), mar = c(4, 4, 3, 1))
on.exit(par(old_par), add = TRUE)

for (i in seq_len(nrow(cells))) {
  hyp <- cells$hypothesis[[i]]
  phase <- cells$phase[[i]]
  precision <- cells$precision[[i]]
  part <- evaluation[
    evaluation$phase == phase & evaluation$precision == precision,
  ]
  part <- part[!duplicated(part$subject_id), ]
  y <- if (hyp == "vram") part$y_vram else part$y_flop_fcm
  t <- if (hyp == "vram") part$target_vram else part$target_flop
  average <- (y + t) / 2
  difference <- y - t
  mean_diff <- mean(difference)
  loa <- 1.96 * sd(difference)

  plot(
    average,
    difference,
    xlab = "(y + t) / 2",
    ylab = "y - t",
    main = paste(hyp, phase, precision),
    pch = 16
  )
  abline(h = 0, lty = 3)
  abline(h = mean_diff, lty = 1)
  abline(h = mean_diff + loa, lty = 2)
  abline(h = mean_diff - loa, lty = 2)
}
```

# Ledger as the sampling frame appendix {#sec-ledger}

Builds that fail, including failures caused by exhausting memory, are recorded
rather than omitted without comment. Those failed attempts do not enter the
eight TOSTs.

```{r}
#| label: ledger
if (nrow(ledger) == 0) {
  cat("No ledger rows.\n")
} else {
  knitr::kable(ledger, caption = "Failed attempts. Primary analysis uses completed subjects only.")
}
```
